PaperMelanoma

### 2022-06-20 Code/PaperMelanoma.Rmd
Limit follow-up to 5 years correctly

# Data preparation
### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# load libraries
library(foreign)
library(mice)
library(survminer)
library(prodlim)
library(rms)
library(Hmisc)
library(png)
set.seed(100)
# import function to create table of missing values
source("Z:/Project Melanoom/PaperMelanoma/Code/table.missing.values.R") 
# import function to plot calibration
source("Z:/Project Melanoom/PaperMelanoma/Code/val_surv_mi_v3.R")
# helper functions for calibration plots
km <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$surv[sf$time==year1]
}
km.lower <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$lower[sf$time==year1]
}
km.upper <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$upper[sf$time==year1]
}
fun.event <- function(lp, h0)
{
  h <- h0*exp(lp)
  p <- 1-exp(-h)
  return(p)
}

@@

### Exclude patients
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc. Note:
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8 (Dbstatus=0: 
negative, Dbstatus=1: positive), even though diagnosis and operation date are 
very close.
```{r, eval=TRUE}
sel.data <- original.data[original.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.data)-nrow(sel.data), "\n")
```
Omit patient A0523 - no last follow-up date
```{r, eval=TRUE}
sel.data[which(is.na(sel.data$Last_FU)), "DatabaseNR"]
sel.data <- sel.data[sel.data$DatabaseNR!="A0523               ",]
```

@@

Omit patients with Rec date before SN date
```{r, eval=TRUE}
old.n <- nrow(sel.data)
sel.data <- sel.data[!(sel.data$Rec.time<0&!is.na(sel.data$Rec.time)),]
cat("Number of patients excluded:", old.n-nrow(sel.data), "\n")
```

@@

#### Delete unused levels
```{r, eval=TRUE}
sel.data <- droplevels(sel.data)
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
data.to.be.imputed <- sel.data[,variables.to.be.imputed]
```
Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd<-datadist(data.to.be.imputed)
options(datadist='dd')
options(digits=8)
```

@@

# Median follow-up time
```{r, eval=TRUE}
median.FU.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~1, data=last.imputed.data.set, reverse=TRUE))
tab.FU.time <- c("total", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                 paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                        sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
median.FU.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~Center, data=last.imputed.data.set, reverse=TRUE))
for (center in sort(as.character(unique(sel.data$Center)))){
  med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
  med.FU <- med.FU.table[3,"quantile"]
  lower.FU <- med.FU.table[3,"lower"]
  upper.FU <- med.FU.table[3,"upper"]
  tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
}
write.table(as.data.frame(tab.FU.time),file='Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt',
            row.names=FALSE)
tab.FU.time
```

# Recurrence at 5 years
```{r, eval=TRUE}
Rec.table <- c("Total", sum(sel.data$Recurrence=="recurred"), sprintf("%.1f", sum(sel.data$Recurrence=="recurred")/nrow(sel.data)*100))
Rec.table <- rbind(Rec.table, c("Positive", 
                                sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Positive"),
                                sprintf("%.1f", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Positive")/sum(sel.data$SNstatus=="Positive")*100)))
Rec.table <- rbind(Rec.table, c("Negative", 
                                sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Negative"),
                                sprintf("%.1f", sum(sel.data$Recurrence=="recurred"&sel.data$SNstatus=="Negative")/sum(sel.data$SNstatus=="Negative")*100)))
utils::write.table(as.data.frame(Rec.table),file='Z:/Project Melanoom/PaperMelanoma/Results/Rec.table.txt',
                   row.names=FALSE)
Rec.table
```

# Melanoma Specific Mortality at 5 years
```{r, eval=TRUE}
MSM.table <- c("Total", sum(sel.data$MSM==1), sprintf("%.1f", sum(sel.data$MSM==1)/nrow(sel.data)*100))
MSM.table <- rbind(MSM.table, c("Positive", sum(sel.data$MSM==1&sel.data$SNstatus=="Positive"),
                                sprintf("%.1f", sum(sel.data$MSM==1&sel.data$SNstatus=="Positive")/sum(sel.data$SNstatus=="Positive")*100)))
MSM.table <- rbind(MSM.table, c("Negative", sum(sel.data$MSM==1&sel.data$SNstatus=="Negative"),
                                sprintf("%.1f", sum(sel.data$MSM==1&sel.data$SNstatus=="Negative")/sum(sel.data$SNstatus=="Negative")*100)))
utils::write.table(as.data.frame(MSM.table),file='Z:/Project Melanoom/PaperMelanoma/Results/MSM.table.txt',
                   row.names=FALSE)
MSM.table
```

# Restrict survival at 5 years
```{r, eval=TRUE}
S.Rec[S.Rec[,1]>5, 1] <- 5 # set recurrence time to 5
S.Rec[S.Rec[,1]>5, 2] <- 0 # set censoring
S.MSM[S.MSM[,1]>5, 1] <- 5 # set follow-up time to 5
S.MSM[S.MSM[,1]>5, 2] <- 0 # set censoring
```

@@

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec) # add a minus to revert class order: from worst to best
  cindex.AJCC.7[i]<-rc["C Index"]
  cindex.AJCC.7.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.7$est),
              paste0("[", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                      "; ", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
            c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.pos$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")),
            c("Negative SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.neg$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.Cindex.txt', row.names=FALSE, col.names=FALSE)

@@

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec) # add a minus to revert class order: from worst to best
  cindex.AJCC.8[i]<-rc["C Index"]
  cindex.AJCC.8.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.8$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                  c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.8.pos$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")),
                  c("Negative SN patients", sprintf('%.2f', cindex.mi.AJCC.8.neg$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.8.Cindex.txt', row.names=FALSE, col.names=FALSE)

# table of number of AJCC.7 and AJCC.8
table.AJCC.7 <- c()
table.AJCC.8 <- c()
for (i in 1:m){
  table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
  table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
}
table.AJCC.7.8 <- cbind(as.numeric(names(table(AJCC.8[,1]))),
                        c(round(rowMeans(table.AJCC.7), 0), 0),
                        round(rowMeans(table.AJCC.8), 0))
utils::write.table(table.AJCC.7.8,
            file='Z:/Project Melanoom/PaperMelanoma/Results/table.AJCC.7.and.8.txt', row.names=FALSE, col.names=FALSE)
table.AJCC.7.8
```

# Full model for recurrence
```{r, eval=TRUE}
form.full.Rec <- S.Rec ~
  SNstatus +
  Sex +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:log(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.Rec <- Hmisc::fit.mult.impute(form.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
f.mi.full.Rec
stats::anova(f.mi.full.Rec)
summary(f.mi.full.Rec)
```

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
c.Breslow <- mean(log(last.imputed.data.set$Breslow))
form.BS.Rec <- S.Rec ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec <- Hmisc::fit.mult.impute(form.BS.Rec,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                           pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
# round(stats::anova(f.mi.BS.Rec)[order(stats::anova(f.mi.BS.Rec)[, 'P']),], 3)
stats::anova(f.mi.BS.Rec)
```

# Refitted full model for melanoma specific mortality
```{r, eval=TRUE}
form.full.MSM.refit <- S.MSM ~
  SNstatus +
  Sex +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:log(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
f.mi.full.MSM.refit <- Hmisc::fit.mult.impute(form.full.MSM.refit,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                                 pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
f.mi.full.MSM.refit
summary(f.mi.full.MSM.refit)
stats::anova(f.mi.full.MSM.refit)
```

>>>>

# Data preparation
### Load packages, functions, and data 
```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# for reproducibility
set.seed(100)

# load libraries
library(foreign)
library(mice)
library(survminer)
library(prodlim)
library(rms)
library(Hmisc)
library(png)
# import function to create table of missing values
source("Z:/Project Melanoom/PaperMelanoma/Code/table.missing.values.R") 
# import function to plot calibration
source("Z:/Project Melanoom/PaperMelanoma/Code/val_surv_mi_v3.R")
# helper functions for calibration plots
km <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$surv[sf$time==year1]
}
km.lower <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$lower[sf$time==year1]
}
km.upper <- function(sel, S.km, horizon){
  S.km.sel <- S.km[sel,]
  sf <- survfit(S.km.sel~1)
  year1 <- max(sf$time[sf$time<=horizon])
  1-sf$upper[sf$time==year1]
}
fun.event <- function(lp, h0)
{
  h <- h0*exp(lp)
  p <- 1-exp(-h)
  return(p)
}

@@

### Inclusion criteria
Exclude when Exclusion = 2: based on duplicates, urogenital melanoma, etc. Note:
Leave the patients included that have EXCLUSIE = 1 and Dbstatus = 8 (Dbstatus=0: 
negative, Dbstatus=1: positive), even though diagnosis and operation date are 
very close.
```{r, eval=TRUE}
sel.data <- original.data[original.data$EXCLUSIE!="Exclusion", selected.vars]
cat("Number of patients excluded:", nrow(original.data)-nrow(sel.data), "\n")
```
Omit patient A0523 - no last follow-up date
```{r, eval=TRUE}
sel.data[which(is.na(sel.data$Last_FU)), "DatabaseNR"]
sel.data <- sel.data[sel.data$DatabaseNR!="A0523               ",]
```

@@

Omit patients with Rec date before SN date (TODO: or impute instead?)
```{r, eval=TRUE}
old.n <- nrow(sel.data)
sel.data[sel.data$Rec.time<0&!is.na(sel.data$Rec.time), "DatabaseNR"]
sel.data <- sel.data[!(sel.data$Rec.time<0&!is.na(sel.data$Rec.time)),]
cat("Number of patients excluded:", old.n-nrow(sel.data), "\n")
```

@@

#### Delete unused levels
```{r, eval=TRUE}
sel.data <- droplevels(sel.data)
save(sel.data,file='Z:/Project Melanoom/PaperMelanoma/Data/sel.data.Rdata', compress=TRUE)
```

# Multiple imputation
```{r, eval=TRUE}
variables.to.be.imputed <- c("Center", "SNstatus", "Sex", "Age.SN", "Ulceration",
                             "Loc_CAT", "Histology", "Breslow", "multiple.fields",
                             "Tot_SNs_neg", "Tot_SNs_pos", "Rdamcrit", "Dewar",
                             "Mitosis",
                             "Recurrence", "dead", "MSM")
data.to.be.imputed <- sel.data[,variables.to.be.imputed]
```

Save distributions of all variables, necessary for Harrell summaries
```{r, eval=TRUE}
dd<-rms::datadist(data.to.be.imputed)
options(datadist='dd')
options(digits=8)
```

@@

# Median follow-up time
```{r, eval=TRUE}
median.FU.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~1, data=last.imputed.data.set, reverse=TRUE))
tab.FU.time <- c("total", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                 paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                        sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
median.FU.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~Center, data=last.imputed.data.set, reverse=TRUE))
for (center in sort(as.character(unique(sel.data$Center)))){
  med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
  med.FU <- med.FU.table[3,"quantile"]
  lower.FU <- med.FU.table[3,"lower"]
  upper.FU <- med.FU.table[3,"upper"]
  tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
}
write.table(as.data.frame(tab.FU.time),file='Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt',
            row.names=FALSE, sep=",")
tab.FU.time
```

# Recurrence at 5 years
```{r, eval=TRUE}
horizon <- 5
SF.Rec.all<-survfit(S.Rec~1)
Rec.all <- 100*(1-SF.Rec.all$surv[SF.Rec.all$time==max(SF.Rec.all$time[SF.Rec.all$time<=horizon])])

SF.Rec.neg<-survfit(S.Rec[sel.data$SNstatus=="Negative"]~1)
Rec.neg <- 100*(1-SF.Rec.neg$surv[SF.Rec.neg$time==max(SF.Rec.neg$time[SF.Rec.neg$time<=horizon])])

SF.Rec.pos<-survfit(S.Rec[sel.data$SNstatus=="Positive"]~1)
Rec.pos <- 100*(1-SF.Rec.pos$surv[SF.Rec.pos$time==max(SF.Rec.pos$time[SF.Rec.pos$time<=horizon])])

SF.MSM.all<-survfit(S.MSM~1)
MSM.all <- 100*(1-SF.MSM.all$surv[SF.MSM.all$time==max(SF.MSM.all$time[SF.MSM.all$time<=horizon])])

SF.MSM.neg<-survfit(S.MSM[sel.data$SNstatus=="Negative"]~1)
MSM.neg <- 100*(1-SF.MSM.neg$surv[SF.MSM.neg$time==max(SF.MSM.neg$time[SF.MSM.neg$time<=horizon])])

SF.MSM.pos<-survfit(S.MSM[sel.data$SNstatus=="Positive"]~1)
MSM.pos <- 100*(1-SF.MSM.pos$surv[SF.MSM.pos$time==max(SF.MSM.pos$time[SF.MSM.pos$time<=horizon])])

Rec.MSM.table <- data.frame(n=c(nrow(sel.data), 
                                sum(sel.data$SNstatus=="Negative"), 
                                sum(sel.data$SNstatus=="Positive")),
                            perc=c(100, 
                                   sprintf("%.1f", sum(sel.data$SNstatus=="Negative")/nrow(sel.data)*100),
                                   sprintf("%.1f", sum(sel.data$SNstatus=="Positive")/nrow(sel.data)*100)),
                            Rec=sprintf("%.1f", c(Rec.all, Rec.neg, Rec.pos)), 
                            MSM=sprintf("%.1f", c(MSM.all, MSM.neg, MSM.pos)))
row.names(Rec.MSM.table) <- c("All", "Negative", "Positive")
utils::write.table(as.data.frame(Rec.MSM.table),file='Z:/Project Melanoom/PaperMelanoma/Results/Rec.MSM.table.txt',
                   row.names=TRUE, sep=",")
Rec.MSM.table
```

# Restrict survival at 5 years
```{r, eval=TRUE}
S.Rec.5 <- S.Rec
S.Rec.5[S.Rec[,1]>5, 1] <- 5 # set recurrence time to 5
S.Rec.5[S.Rec[,1]>5, 2] <- 0 # set censoring

S.MSM.5 <- S.MSM
S.MSM.5[S.MSM[,1]>5, 1] <- 5 # set follow-up time to 5
S.MSM.5[S.MSM[,1]>5, 2] <- 0 # set censoring
```

@@

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.7.i, S.Rec.5) # add a minus to revert class order: from worst to best
  cindex.AJCC.7[i]<-rc["C Index"]
  cindex.AJCC.7.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.7.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.7.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.7.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.7.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.7.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.7<-Rubin.combine(cindex.AJCC.7,cindex.AJCC.7.se)
cindex.mi.AJCC.7.pos<-Rubin.combine(cindex.AJCC.7.pos,cindex.AJCC.7.se.pos)
cindex.mi.AJCC.7.neg<-Rubin.combine(cindex.AJCC.7.neg,cindex.AJCC.7.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.7$est),
              paste0("[", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.025)*cindex.mi.AJCC.7$se),
                      "; ", sprintf('%.2f', cindex.mi.AJCC.7$est+qnorm(.975)*cindex.mi.AJCC.7$se), "]")),
            c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.pos$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.025)*cindex.mi.AJCC.7.pos$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.pos$est+qnorm(.975)*cindex.mi.AJCC.7.pos$se), "]")),
            c("Negative SN patients:", sprintf('%.2f', cindex.mi.AJCC.7.neg$est),
                  paste0("[", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.025)*cindex.mi.AJCC.7.neg$se),
                         "; ", sprintf('%.2f', cindex.mi.AJCC.7.neg$est+qnorm(.975)*cindex.mi.AJCC.7.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.7.Cindex.txt', 
            row.names=FALSE, col.names=FALSE, sep=",")

@@

  # compute C-index
  rc <- Hmisc::rcorr.cens(-AJCC.8.i, S.Rec.5) # add a minus to revert class order: from worst to best
  cindex.AJCC.8[i]<-rc["C Index"]
  cindex.AJCC.8.se[i]<-rc["S.D."]/2

  # compute C-index for only positive sentinel node patients
  rc.pos <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Positive"], S.Rec.5[imputation.i$SNstatus=="Positive",])
  cindex.AJCC.8.pos[i]<-rc.pos["C Index"]
  cindex.AJCC.8.se.pos[i]<-rc.pos["S.D."]/2

  # compute C-index for only negative sentinel node patients
  rc.neg <- Hmisc::rcorr.cens(-AJCC.8.i[imputation.i$SNstatus=="Negative"], S.Rec.5[imputation.i$SNstatus=="Negative",])
  cindex.AJCC.8.neg[i]<-rc.neg["C Index"]
  cindex.AJCC.8.se.neg[i]<-rc.neg["S.D."]/2
}
# combine C-index for each imputed data set
cindex.mi.AJCC.8<-Rubin.combine(cindex.AJCC.8,cindex.AJCC.8.se)
cindex.mi.AJCC.8.pos<-Rubin.combine(cindex.AJCC.8.pos,cindex.AJCC.8.se.pos)
cindex.mi.AJCC.8.neg<-Rubin.combine(cindex.AJCC.8.neg,cindex.AJCC.8.se.neg)

# save C-index and confidence interval
utils::write.table(rbind(c("All patients:", sprintf('%.2f', cindex.mi.AJCC.8$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.025)*cindex.mi.AJCC.8$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8$est+qnorm(.975)*cindex.mi.AJCC.8$se), "]")),
                  c("Positive SN patients:", sprintf('%.2f', cindex.mi.AJCC.8.pos$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.025)*cindex.mi.AJCC.8.pos$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.pos$est+qnorm(.975)*cindex.mi.AJCC.8.pos$se), "]")),
                  c("Negative SN patients", sprintf('%.2f', cindex.mi.AJCC.8.neg$est),
                        paste0("[", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.025)*cindex.mi.AJCC.8.neg$se),
                               "; ", sprintf('%.2f', cindex.mi.AJCC.8.neg$est+qnorm(.975)*cindex.mi.AJCC.8.neg$se), "]"))),
            file='Z:/Project Melanoom/PaperMelanoma/Results/AJCC.8.Cindex.txt', 
            row.names=FALSE, col.names=FALSE, sep=",")

# table of number of AJCC.7 and AJCC.8
table.AJCC.7 <- c()
table.AJCC.8 <- c()
for (i in 1:m){
  table.AJCC.7 <- cbind(table.AJCC.7, as.numeric(table(AJCC.7[,i])))
  table.AJCC.8 <- cbind(table.AJCC.8, as.numeric(table(AJCC.8[,i])))
}
table.AJCC.7.8 <- cbind(as.numeric(names(table(AJCC.8[,1]))),
                        c(round(rowMeans(table.AJCC.7), 0), 0),
                        round(rowMeans(table.AJCC.8), 0))
utils::write.table(table.AJCC.7.8,
            file='Z:/Project Melanoom/PaperMelanoma/Results/table.AJCC.7.and.8.txt', 
            row.names=FALSE, col.names=FALSE, sep=",")
table.AJCC.7.8
```

# Test non-lineairities
### Test if Age should be modelled non-linearly
```{r, eval=TRUE}
form.linear.full.Rec <- S.Rec.5 ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
form.nonlinear.age.full.Rec <- S.Rec.5 ~
  SNstatus +
  Sex +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:log(Age.SN) +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
linear.full.Rec <- Hmisc::fit.mult.impute(form.linear.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
nonlinear.age.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.age.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
# distribution of Age
hist(last.imputed.data.set$Age.SN)
# linear model
linear.full.Rec$loglik[2]
# using log(Age) gives slightly better log likelihood
nonlinear.age.full.Rec$loglik[2]
```

### Test if Breslow should be modelled non-linearly
```{r, eval=TRUE}
form.nonlinear.Breslow.full.Rec <- S.Rec.5 ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  log(Breslow) +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  Rdamcrit +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:log(Breslow) +
  SNstatus:multiple.fields
nonlinear.Breslow.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.Breslow.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
# distribution of Breslow
hist(last.imputed.data.set$Breslow)
# linear model
linear.full.Rec$loglik[2]
# using log(Breslow) gives better log likelihood
nonlinear.Breslow.full.Rec$loglik[2]
```

### Test if Rdamcrit should be modelled non-linearly
```{r, eval=TRUE}
form.nonlinear.Rdamcrit.full.Rec <- S.Rec.5 ~
  SNstatus +
  Sex +
  Age.SN +
  Ulceration +
  Loc_CAT +
  Histology +
  Breslow +
  multiple.fields +
  Tot_SNs_neg +
  Tot_SNs_pos +
  log(Rdamcrit) +
  Dewar +
  SNstatus:Sex +
  SNstatus:Age.SN +
  SNstatus:Ulceration +
  SNstatus:Loc_CAT +
  SNstatus:Histology +
  SNstatus:Breslow +
  SNstatus:multiple.fields
nonlinear.Rdamcrit.full.Rec <- Hmisc::fit.mult.impute(form.nonlinear.Rdamcrit.full.Rec, cph, xtrans=imputed.data,
                                 data=data.to.be.imputed,n.impute=m, pr=FALSE,
                                 fit.reps=TRUE, y=TRUE, x=TRUE, se.fit=TRUE)
# distribution of Rdamcrit
hist(last.imputed.data.set$Rdamcrit)
# linear model
linear.full.Rec$loglik[2]
# using log(Rdamcrit) gives better log likelihood
nonlinear.Rdamcrit.full.Rec$loglik[2]
```
